Supplementary Material for 'The complex network of trophic interactions in

### 2023-07-11 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Schejter suggs.

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, 53º–55ºS and 56º–62ºW, respectively), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1788 predator-prey interactions and 380 species, with a link density of 4.70 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, such as the South Georgia and South Sandwich, Bouvet, Prince Edward, and Macquarie islands [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, 53º–55ºS 56º–62ºW), which is the focus of this work, is unique since these MPAs are associated to a submarine plateau and its southern adjacent deep slope region, respectively [@Falabella2017]. In addition, such MPAs are part of a network of protected areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to protect this southern region in order to contribute to global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers (or proportions) of endemic species, and/or high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a potentially sensitive and biodiverse benthic habitat that was only barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by high biomass of vulnerable and fragile species (mainly Porifera, Bryozoa and Cnidaria) that considered with their environment meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResourcesCCAMLR2009]. Also, the benthic realm provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Vazquez2018; @Delpiani2020; @Troccoli2020; @CovattiAle2022; @Fischer2022; @Matusevich2022; @Florencia2023]. The maintenance of this singular community is related to local and regional oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020].

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws 26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB II, respectively [@Tombesi2020; @AdministraciondeParquesNacionales2022] (Figure 1). They are located 150 km east of Isla de los Estados and 200 km south of Malvinas/Falkland Islands.
The MPA N-BB I comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. The N-BB II includes such a deep slope, protecting about 32,000 km^2^. Intense upwelling and mixing occur in relation with the slope, entraining deep nutrient-rich waters into the photic layer [@Piola2009; @Matano2019] and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

@@

load("../results/summary_results_jun23.rda")

@@

load("../data/foodweb-data_jun23.rda")
load("../results/summary_results_jun23.rda")

@@

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). The most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight those high-trophic level species (TL > 3.1) with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Daption capense* and *Eudyptes chrysocome* (seabirds), Ziphiidae and *Lagenorhynchus cruciger* (marine mammals) (Table S3).

@@

We have identified that most of the consumers in the food web of the MPA N-BB either have a narrow diet or are specialists, while few present a broad or generalist diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed in the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources on which most consumers depend and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016], likely feeding on detritus that is constantly resuspended from the bottom [@Martin2021]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators such as the black-browed and grey-headed albatrosses (*Thalassarche melanophris* and *Thalassarche chrysostoma*, respectively) [@Catry2004], and baleen whales (species of the genera Balaenoptera and Eubalaena) [@Valenzuela2018].

@@

Closeness and betweenness are defined as "mesoscale" properties because they consider direct and indirect interactions, therefore describing the focal species’ ability to influence the rest of the species of the food web [@Lai2012].
Closeness quantifies how many steps away species $i$ is from all other species in the food web, and is proportional to how rapidly the indirect effects of the focal species can spread to other species in the network [@Scotti2010]. In the food web of the MPA N-BB, low-TL consumers arise as important in this regard: species of the zooplankton community, *Calanus* and *Euphausia*, and Brachiopoda. Any perturbation affecting these species, such as the recently confirmed contaminants mercury [@Fioramonti2022] and microplastics [@Cossi2021; @DiMauro2022], should be of concern since it might reach many other species in the food web.
Otherwise, betweenness measures the number of shortest paths between species, providing information on the importance of species as “bridges” for energy transfer: a species with high betweenness takes part in more food chains and therefore affects more energy flows [@Scotti2010]. We have identified the longtail southern cod *Patagonotothen ramsayi* as the most important species in this sense. Moreover, in light of our analysis, other species like the Patagonian toothfish *Dissostichus eleginoides*, the Patagonian cod *Salilota australis*, the yellowfin notothenioid *Patagonotothen guntheri*, and the Patagonian longfin squid *Doryteuthis gahi* should be considered as relevant in the energy transfer in the ecosystem. All these species have a mid-trophic position in the food web, supporting the 'wasp-waist' control hypothesis for the MPA N-BB [@Riccialdelli2020].

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. The significant negative relationship between trophic similarity and trophic level enhances the hypothesis of functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web [@Riccialdelli2020]. At the same time, our results highlight the uniqueness in terms of the trophic role of high-TL predators. Here, not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also demersal vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and benthic invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level. In this regard, we should mention the potential threat of the fisheries operating in the western section of the MPA N-BB, where this activity is allowed and mostly focuses on the Patagonian toothfish *Dissostichus eleginoides* and the southern blue whiting *Micromesistius australis* [@Martinez2015]. Although the fishing effort is concentrated outside the limits of the MPA N-BB, the impact on the MPA ecosystem should not be neglected [@Martinez2021].

@@

It's important to mark that we did not consider quantitative data (i.e. abundance, biomass) to assess the species' role in the food web. Although there exists such data for some species [@Schejter2021], it would not be possible to include it in the food web framework described here due to a taxonomical resolution mismatch. In this regard, we should mention the case of *Zygochlamys patagonica* (Bivalvia) that is highlighted by our species-level analyses though it has been found in low abundances in the area [@Schejter2021].

>>>>

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1788 predator-prey interactions and 380 species, with a link density of 4.70 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, such as the South Georgias and South Sandwich, Bouvet, Prince Edward, and Macquarie islands [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS ~56º–62ºW), which is the focus of this work, is unique since these MPAs are associated to a submarine plateau and its southern adjacent deep slope region, respectively [@Falabella2017; @Schejter2020]. In addition, such MPAs are part of a network of protected areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to protect this southern region in order to contribute to global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers (or proportions) of endemic species, and/or high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a potentially sensitive and biodiverse benthic habitat that was only barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by high biomass of vulnerable and fragile species (mainly Porifera, Bryozoa and Cnidaria) that considered with their environment meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present densities of Indicator Taxa of > 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResourcesCCAMLR2009]. Also, the benthic realm provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Vazquez2018; @Delpiani2020; @Troccoli2020; @CovattiAle2022; @Fischer2022; @Matusevich2022; @Florencia2023]. The maintenance of this singular community is related to local and regional oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020].

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws 26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB II, respectively [@Tombesi2020; @AdministraciondeParquesNacionales2022] (Figure 1). They are located 150 km east of Isla de los Estados and 200 km south of Malvinas/Falkland Islands.
The MPA N-BB I comprises nearly 28,900 km^2^ circumscribed by the 200 m isobath, between ~54º–55ºS and ~56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. The N-BB II includes such a deep slope, protecting about 32,000 km^2^ (~55º-56ºS, ~58º-62ºW). Intense upwelling and mixing occur in relation with the slope, entraining deep nutrient-rich waters into the photic layer [@Piola2009; @Matano2019] and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

@@

load("../results/summary_results_jul23.rda")

@@

load("../data/foodweb-data_jul23.rda")
load("../results/summary_results_jul23.rda")

@@

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). The most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight those high-trophic level species (TL > 3.1) with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster* sp. (echinoderms), *Daption capense* and *Eudyptes chrysocome* (seabirds), Ziphiidae and *Lagenorhynchus cruciger* (marine mammals) (Table S3).

@@

We have identified that most of the consumers in the food web of the MPA N-BB either have a narrow diet or are specialists, while few present a broad or generalist diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], juveniles of *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (with the exception of adults of *Dissostichus eleginoides* that are top predators), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed in the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources on which most consumers depend and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016], components of the animal forest [@Schejter2020], likely feeding on detritus that is constantly resuspended from the bottom [@Martin2021]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators such as the black-browed and grey-headed albatrosses (*Thalassarche melanophris* and *Thalassarche chrysostoma*, respectively) [@Catry2004], and baleen whales (species of the genera *Balaenoptera* and *Eubalaena*) [@Valenzuela2018].

@@

Closeness and betweenness are defined as "mesoscale" properties because they consider direct and indirect interactions, therefore describing the focal species’ ability to influence the rest of the species of the food web [@Lai2012].
Closeness quantifies how many steps away species $i$ is from all other species in the food web, and is proportional to how rapidly the indirect effects of the focal species can spread to other species in the network [@Scotti2010]. In the food web of the MPA N-BB, low-TL consumers arise as important in this regard: species of the zooplankton community, *Calanus* and *Euphausia*, *Zygochlamys patagonica* (Bivalvia), and Brachiopoda. Any perturbation affecting these species, such as the recently confirmed contaminants mercury [@Fioramonti2022] and microplastics [@Cossi2021; @DiMauro2022], should be of concern since it might reach many other species in the food web.
Otherwise, betweenness measures the number of shortest paths between species, providing information on the importance of species as “bridges” for energy transfer: a species with high betweenness takes part in more food chains and therefore affects more energy flows [@Scotti2010]. We have identified the longtail southern cod *Patagonotothen ramsayi* as the most important species in this sense. Moreover, in light of our analysis, other species like the Patagonian toothfish *Dissostichus eleginoides* (juveniles), the Patagonian cod *Salilota australis*, the yellowfin notothenioid *Patagonotothen guntheri*, and the Patagonian longfin squid *Doryteuthis gahi* should be considered as relevant in the energy transfer in the ecosystem. All these species have a mid-trophic position in the food web, supporting the 'wasp-waist' control hypothesis for the MPA N-BB [@Riccialdelli2020].

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. The significant negative relationship between trophic similarity and trophic level enhances the hypothesis of functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web [@Riccialdelli2020]. At the same time, our results highlight the uniqueness in terms of the trophic role of high-TL predators. Here, not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also demersal vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and benthic invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster* sp.) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level. In this regard, we should mention the potential threat of the fisheries operating in the western section of the MPA N-BB, where this activity is allowed and mostly focuses on the Patagonian toothfish *Dissostichus eleginoides* and the southern blue whiting *Micromesistius australis* [@Martinez2015]. Although the fishing effort is concentrated outside the limits of the MPA N-BB, the impact on the MPA ecosystem should not be neglected [@Martinez2021].

@@

It's important to mark that we did not consider quantitative data (i.e. abundance, biomass) to assess the species' role in the food web. Although there exists such data for some species [@Schejter2021], it would not be possible to include it in the food web framework described here due to a taxonomical resolution mismatch. In this regard, we should mention the case of *Zygochlamys patagonica* (Bivalvia) and Brachiopoda that are highlighted by our species-level analyses though they have been found in low abundances in the area [@Schejter2021].

### 2023-07-11 MS_ToSubmit/SupMaterial.Rmd
Schejter suggs.

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  dplyr::select(Prey, Predator, Reference, Reference_link) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

>>>>

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  mutate(Prey = case_when(Prey == "Brachiopoda_spp" ~ "Brachiopoda", TRUE ~ "Other")) %>%
  dplyr::select(Prey, Predator, Reference, Reference_link) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)
